---
title: "a3_task3_renee_lamanna"
author: "Renee LaManna"
date: "10/27/2021"
output: html_document
---

## Setup

- Attatching packages
- Set codechunk options

```{r setup, include = TRUE, warnings = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)

library (tidyverse)
library (here)
library (janitor)
```

## Read in Data

- urchins.csv = file containing data for red and purple urchins observed at 5 sites along the Santa Barbara coast

```{r}
# Read in urchins.csv as `urchins`
urchins <- read_csv(here("data", "urchins.csv"))
```

## Data wrangling and cleaning

- Using clean_names(), create a new dataset, `urchins_clean` that is converted to lowersnakecase

```{r}
urchins_clean <- urchins %>% 
  clean_names()
```

## Data Visualization

- Create histogram comparing distribution of sizes between purple and red urchins
- Use urchins_clean as data and indicate size on the x-axis, the y-axis will automatically be count
- Use geom_historgam() to indicate type of graph and set bin size to 10
- Use facet_wrap() fxn to organize purple and red urchins into separate graphs
- Added theme_minimal(), renames axis, added title, subtitle, and caption all with labs()

```{r}
ggplot(data = urchins_clean, aes(x = size)) +
  geom_histogram(bins = 10) +
  facet_wrap(~ common_name) +
  theme_minimal() +
  labs( x = "Urchin diameter (cm)", 
        y = "Number of urchins",
        title = "Urchin diameter distribution for purple and red urchins",
        subtitle = "Santa Barbara, CA",
        caption = "Figure 1. Size, represented as urchin diameter (cm) distribution for both purple and red urchins at 5 different sites along Santa Barbara coast")
```



