---
title: "Task 4: Purple Urchins at Naples Reef w/ and w/out Kelp Removal"
author: "Renee LaManna"
date: "10/28/2021"
output: html_document
---
## Overview 
This code wrangles data to compare size distributions of purple urchins under control treatments where no kelp was removed and annual treatments where kelp was removed at Naples Reef. The data is then visualized with an exploratory histogram and QQ-plot to determine normality. Finally, a summary statistics table is also created to compare the size distribution of purple urchins between the controlled and annual treatments. This data came from a scientific article titled, "SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution", written by Daniel Reed and published in 2020. The dataset, urchins.csv describes the size frequency distribution of red and purple sea urchins at Santa Barbara Channel - Long Term Ecological Research long-term kelp removal experiment sites.

Data Citation: Santa Barbara Coastal LTER and D. Reed. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative. https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b

## Setup
Attaching packages and set codechunk options

```{r setup, include = TRUE, warnings = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(here)
library(janitor)
library(lubridate)
library(kableExtra)
```

## Read in Data
read in urchins.csv, file containing data for red and purple urchins observed at 5 sites along the Santa Barbara coast

```{r}
urchins <- read_csv(here("data", "urchins.csv"))
```


## Data Wrangling
Wrangle data to only include Naples Reef observations for purple urchins observed in 2010

```{r}
naples_urchins <- urchins %>% 
  clean_names() %>% 
  filter(common_name == "Purple Urchin") %>% 
  mutate(date = mdy(date)) %>% 
  mutate(year = year(date)) %>% 
  filter(year == 2010) %>% 
  filter(site == "NAPL") 
  
```

## Data Visualization
Made exploratory faceted histograms and qq-plots comparing annual(kelp removal) vs control (no kelp removal) treatments to determine if data is normally distributed

```{r}
# Make exploratory histogram
ggplot(data = naples_urchins, aes(x = size)) +
  geom_histogram(bins = 12) +
  facet_wrap(~treatment)

# Make exploratory QQ-plot
ggplot(data = naples_urchins, aes(sample = size)) +
  stat_qq() +
  geom_qq_line() + # plot line to compare linear relationship
  facet_wrap(~ treatment)
```

Given the exploratory faceted histograms and QQ-plots above, I would consider the size distribution of purple urchins at Naples in 2010 both in control and annual to be normally distributed. The QQ-plot shows a somewhat linear pattern. I added a linear line for better comparison. Most of the observations appear to fall within that line, especially within the annual treatment, with a few less in the control. Additionally, the histogram is approximately bell-shaped and symmetrical. After visual observation of the faceted histograms and QQ-plots, I assume normality.

## Summary Statistics
Create a summary statistic table with the mean, median, sd, and count of urchin size grouping by treatment

```{r}
# Data wrangling to prepare for table
naples_summary <- naples_urchins %>% 
  group_by(treatment) %>% 
  summarise(mean = round(mean(size), 2), 
            median = median(size), 
            sd = round(sd(size), 2), 
            count = n())

# kable() from the kableExtra package used to finalize table
kable(naples_summary,
      col.names = c("Species", "Mean (cm)", "Median (cm)", "SD (cm)", "Count"),
      caption = "Table 1: Mean, median, sd, and count values of purple urchins' diameter(cm) under controlled and annual treatments at Naples Reef, Santa Barbara", 
      digit = 2) %>% 
      kable_styling(full_width = FALSE,
                bootstrap_options = "bordered",
                position = "left") 
  
```


## End Task

