---
title: "Task 3: Size Frequency Distribution of Urchins at SBC-LTER "
author: "Renee LaManna"
date: "10/27/2021"
output: html_document
---

## Overview 
This code wrangles and visualizes data to compare size distributions of purple and red urchins along the Santa Barbara coast in Section 1. A summary statistics table is also created for this same data. Then, in Section 2, this code wrangles and visualizes data exploring purple urchin counts by site and year. This data came from a scientific article titled, "SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution", written by Daniel Reed and published in 2020. The dataset, urchins.csv is used for both sections and describes the size frequency distribution of red and purple sea urchins at Santa Barbara Channel - Long Term Ecological Research long-term kelp removal experiment sites.

Data Citation: Santa Barbara Coastal LTER and D. Reed. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative. https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b 

## Setup
Attach packages and set codechunk options

```{r setup, include = TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library (tidyverse)
library (here)
library (janitor)
library (kableExtra)
library (DT)
library (lubridate)

```

## Read in Data
 urchins.csv = file containing data for red and purple urchins observed at 5 sites along the Santa Barbara coast

```{r}
# Read in urchins.csv as `urchins`
urchins <- read_csv(here("data", "urchins.csv"))
```

## Data Cleaning
Using clean_names(), create a new dataset, `urchins_clean` that is converted to lowersnakecase

```{r}
urchins_clean <- urchins %>% 
  clean_names()
```

# Section 1:

## Data Visualization
Create histogram comparing distribution of sizes between purple and red urchins using ggplot

```{r}
# Plot and change formatting to be clear and attractive
ggplot(data = urchins_clean, aes(x = size)) +
  geom_histogram(aes(fill = common_name), bins = 10) +
  scale_fill_manual(values = c("mediumorchid4", "red4")) +
  facet_wrap(~common_name) +
  theme_minimal() +
  theme(axis.title.x = element_text(hjust = 0.5, vjust = 0.5),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        plot.caption = element_text(hjust = 0.05),
        legend.position = "none") +
  labs( x = "Urchin diameter (cm)", 
        y = "Number of urchins",
        title = "Urchin diameter distribution for purple and red urchins",
        subtitle = "Santa Barbara, CA",
        caption = "Figure 1. Size distribution, represented as urchin diameter (cm) for both purple and red urchins at 5 different sites 
        along Santa Barbara coast")
```


## Summary Statistics 
Create finalized summary table that contains the following statistics for purple and red urchins: mean, median, standard deviation,and sample size

```{r}
# Data wrangling to prepare for table
urchins_summary <- urchins_clean %>% 
  group_by(common_name) %>% # grouping variable is species
  summarise(mean = mean(size),  
            median = median(size), 
            sd = sd(size), 
            count = n())

# kable() from the kableExtra package used to finalize table 
kable(urchins_summary, 
      col.names = c("Species", "Mean (cm)", "Median (cm)", "SD (cm)", "Count"),
      caption = "Table 1: Mean, median, sd, and count values for purple and red urchins at SBC-LTER sites", 
      digit = 2) %>% 
      kable_styling(full_width = FALSE,
                bootstrap_options = "bordered",
                position = "left")
         
  
```

The red urchin had a higher mean, median, and SD when compared to the purple urchin. The actual mean difference in urchin diameter between the two is 2.31 cm. The percent mean difference is 41.3%. Their central tendency also differs by 2.5. Their overall spread has a difference of 0.71. 


# Section 2: 

## Data Wrangling
Wrangle data to only contain observations for purple urchins, using mutate(), convert date to class "date", create a year column, and a column with full site names

```{r}
purple_urchins <- urchins_clean %>% 
  filter(common_name == "Purple Urchin") %>% 
  mutate(date = mdy(date)) %>% 
  mutate(year = year(date)) %>% 
  mutate(site_full = case_when(site == "NAPL" ~ "Naples Reef",
                               site == "AQUE" ~ "Arroyo Quemado",
                               site == "CARP" ~ "Carpinteria Reef",
                               site == "MOHK" ~ "Mohawk Reef",
                               site == "IVEE" ~ "Isla Vista Reef"))
```

Then create new subset that groups count of purple urchins by site and year

```{r}
total_purple_urchins <- purple_urchins %>% 
  group_by(site_full, year) %>% 
  summarise(count=n())
```


## Data Visualization
Make final graph of pruple urchins recorded at each site over time. Included both a column graph and line graph.

```{r}
ggplot(data = total_purple_urchins, aes(x = year, y = count)) + 
  geom_col(aes(color = site_full, fill = site_full)) +
  facet_wrap(~site_full, scales = "free") +
  scale_color_manual(values = c("white", "white", "white", "white", "white")) +
  scale_fill_manual(values = c("light blue", "seagreen", "khaki2", "lightsalmon3", "orangered4")) +
  theme_gray() +
  theme(axis.title.x = element_text(hjust = 0.5, vjust = 0.5),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 60, size = 6),
        plot.caption = element_text(hjust = .4),
        legend.position = "none") +
  scale_x_continuous(limits=c(2007,2020)) + 
  scale_y_continuous(limits=c(0,1200)) +
  labs( x = "Year", 
        y = "Number of urchins",
        title = "Number of Purple Urchins per Site (LAMANNA)", 
        subtitle = "SBC-LTER: 2008-2020",
        caption = "Figure 2. Size distribution, represented as urchin diameter (cm) for both purple urchins on Santa Barbara coast")
```
```{r}
ggplot(data = total_purple_urchins, aes(x = year, y = count)) + 
  geom_line(aes(color = site_full)) +
  scale_color_manual(values = c("light blue", "seagreen", "khaki2", "lightsalmon3", "orangered4")) +
  theme_gray() +
  theme(axis.title.x = element_text(hjust = 0.5, vjust = 0.5),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        axis.text.x = element_text(size = 6),
        plot.caption = element_text(hjust = .4),
        legend.position = "right") +
  scale_x_continuous(limits=c(2007,2020)) + 
  scale_y_continuous(limits=c(0,1200)) +
  labs( x = "Year", 
        y = "Number of urchins",
        title = "Number of Purple Urchins per Site (LAMANNA)", 
        subtitle = "SBC-LTER: 2008-2020",
        caption = "Figure 2. Size distribution, represented as urchin diameter (cm) for both purple urchins on Santa Barbara coast",
        color = "Site Location",
        Key = "Site")
```

I created two graphs, a histogram and a line for improved comparison. I believe the graphs reflect a decrease in purple urchin population. However here we included observations both from control and annual. Therefore, the data doesn't truly represent "actual changes". So in order to determine that the changes in counts per year are reflective of actual changes, we could separate the control from the annual. Changes in urchin population can also be due to sampling error or other environmental changes besides kelp removal such as warming sea surface temperatures. 


## End Task